### 2014-10-21 Publications/2013-TMI/05-experiments.tex
updated methods

% -*- root: 00-main.tex -*-
\section{Data and experiments}
\label{sec:experiments}
%

\subsection{Image Data}
\label{sec:datasets}

\subsubsection{Simulated digital phantom} %
%
The lack of a widely accepted \emph{gold}-standard in the application
field has been addressed by several authors \citep{cote_tractometer_2013}
In this work, we use one of the most complete, and publicly available,
digital phantoms (\url{http://hardi.epfl.ch/static/events/2013_ISBI/testing_data.html}).
The phantom is a spherical volume containing a set of fiber bundles, that connect
one area of a ``cortex'' to another. The model accounts for \gls{pve} using
a similar approach to \citep{close_software_2009} and for \gls{csf} contamination
as well.

\subsubsection{Real \gls{mri} datasets} %
%
We used image data from XX healthy volunteers with no history of neurological
conditions (ages XX$\pm$XX, X female) to illustrate the applicability of our
approach. All the subjects were scanned in a 3T MR Scanner (Siemens Magnetom
TrioTim, Siemens, Erlangen, Germany) with a standard 12-channel head coil.
\todo[inline]{check coil \# channels}
Subjects were scanned twice with the same protocol,
described hereafter. After being scanned the first time, each subject exited the
scan room for a short break and then reentered for an identical scan session.
To note, there was a full repositioning of the volunteer, coils, blankets and pads
before each scan and re-scan session. The scan session protocol was as follows:
\begin{enumerate}
\item Triplanar survey (Localizer).
\item Field Mapping: field mapping using GRE sequence was performed before \gls{dwi}
acquisition for susceptibility correction purposes.
\item DTI: \gls{dwi} were acquired with axial in-plane
isotropic resolution $2mm$, slice thickness $2mm$, $XXX \times XXX \times XXX$
image matrix, TR= XXXX ms, TE = XX ms, NEX = X, BW = XXXX Hz/pixel, GRAPPA
acceleration factor 2. The series included images acquired with diffusion
weighting along 30 non-collinear directions ( $b$=700sm$^{-2}$ ) sampled
twice for averaging, and 5 interleaved images acquired without diffusion
weighting ( $b=0$ ).
\item Field Mapping (same as before DTI)
\item Structural T1: An MPRAGE T1-weighted acquisition, sagittal
GRE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
$XXX \times XXX \times XXX$ image matrix, TR=2300 ms, TE=2.98 ms, FA= 9, NEX=X, BW= 240 Hz/pixel.
\item Structural T2: A T2-weighted acquisition, oblique axial
TSE sequence, in-plane isotropic resolution 1.0 mm, slice thickness 1.2mm
$XXX \times XXX \times XXX$ image matrix, TR=3200 ms, TE= 408 ms, NEX=X, BW= 751 Hz/pixel.
\end{enumerate}

\subsection{Image preprocessing and shape-prior generation}
\label{sec:preprocessing}
%
Regardless the dataset type (simulated or real), all \gls{dwi} datasets were processed
using the standard \gls{dti} reconstruction methods provided by FSL \footnote{DTIFIT,
included in the FMRIB's Software Library (FSL),
\url{http://fsl.fmrib.ox.ac.uk/fsl/fsl-4.1.9/fdt/fdt_dtifit.html}} to fit
the tensor model and produce scalar maps of the required features.
The properties of the reconstructed tensors and derived scalar maps have
been studied by \cite{ennis_orthogonal_2006}. Based on their
findings, \gls{fa}~\eqref{eq:fa} and \gls{md}~\eqref{eq:md} are
considered complementary features, and therefore we selected them for the
energy model \eqref{eq:complete_energy} in driving the
registration-segmentation process.
Whereas \gls{fa} informs mainly about the \emph{shape} of diffusion,
the \gls{md} is more related to the \emph{magnitude} of the process:

\begin{align}
\mathrm{FA} &= \sqrt{ \frac{1}{2}}\,\frac{\sqrt{ (\lambda_1 - \lambda_2)^2 + (\lambda_2 - \lambda_3)^2 + (\lambda_3 - \lambda_1)^2}}{\sqrt{ {\lambda_1}^2 + {\lambda_2}^2 + {\lambda_3}^2}} \label{eq:fa} \\
\mathrm{MD} &= ( \lambda_1 + \lambda_2 + \lambda_3 ) / 3 \label{eq:md}
\end{align}
where $\lambda_i$ are the eigenvalues of the diffusion tensor
associated with the diffusion signal $S(\vec{q})$. There exist
two main reasons to justify their choice.
First, they are well-understood and standardized in clinical routine.
Second, together they contain most of the information that is
usually extracted from the \gls{dwi}-derived scalar maps
\cite{ennis_orthogonal_2006}.

Preprocessing differed between simulated and real datasets in the
shape-prior surfaces generation, as we describe hereafter.

@@

\subsection{Experiments and evaluation}
\label{sec:experiments_evaluation}
%
\subsubsection{Validation on the simulated digital phantom} %
\label{sec:validation_phantom}
We firstly evaluated our approach on the simulated data, using
the distorted data with a deformation field similar to the
susceptibility artifact that affects real \gls{dwi} data. To
this end, we report the following indices:
\begin{itemize}
\item \gls{se}, a distance between the one-to-one corresponding
vertices, weighted by their respective Voronoi area.
\begin{equation}
MSE = \sum\limits_k \sum\limits_j^M w_j\,\| \mathbf{x}_j - \hat{\mathbf{x}}_j \|
\end{equation}
where $\mathbf{x}_j$ are the locations of the $M$ vertices of the $k$ priors,
$\hat{\mathbf{x}}_j$ are the corresponding locations recovered, and $w_j$ the
weighting factor as the relative surface of the Voronoi area.
\item \gls{wi}, L2-distance between the theoretical and the recovered
deformation field.
\begin{equation}
WI = \frac{1}{N} \sum\limits_i^N \| \mathbf{d}_i - \hat{\mathbf{d}}_i \|
\end{equation}
where $\mathbf{d}_i$ is the theoretical displacement vector at position $i$
and $\hat{\mathbf{d}}_i$ is the recovered one at the same index position.
\item Parcellation agreement, the \gls{se} averaged by defined \glspl{roi}
between the theoretical and the recovered parcellations.
\item \gls{nof} agreement, between the fibers recovered on the original
data and the processed data.
\end{itemize}

Additionally, the entropy of data is studied in both
original and recovered datasets to draw another possible basis for the
assessment of the real datasets. We also report the same indices for the
outcome of a widely-used methodology that combines field-map susceptibility
correction and T1-T2-\gls{dwi} registration .
\todo[inline]{I suggest to give it a compact name and define it in the
pre-processing section or even before in the intro}

\subsubsection{Evaluation on real datasets}
For the real datasets there is no published \emph{gold}-standard to validate results.
Thus, visual results are provided to let compare the performance with the NICEACRONYM
standard methodology. Additionally, cross-comparison of repeatability results are
provided. In this second evaluation strategy, all the indices defined in
\autoref{sec:validation_phantom} are reported.

>>>>

% -*- root: 00-main.tex -*-
\section{Data and experiments}
\label{sec:experiments}

@@

### 2014-10-21 Publications/2013-TMI/08-appendix-01.tex
updated methods

\appendices

@@

>>>>

\appendix

@@

{\color{red} {Using the region descriptors derived in \autoref{sec:methods_map}, we propose
an \gls{adf}-like, \gls{acwe}-based, piece-wise constant, image segmentation
model (where the unknown is the deformation field)
\cite{chan_active_2001} with the energy functional obtained in
\eqref{eq:final_map_energy}. This inverse problem is ill-posed
\cite{bertero_illposed_1988,hadamard_sur_1902}.
In order to account for deformation field regularity and to render the
problem well-posed, we include limiting and regularization terms into
the energy functional \cite{morozov_linear_1975,tichonov_solution_1963}:
%
\begin{multline}
E(u) = \sum\limits_l \int_{\Omega'_l} \mdist{f}{l} \,d\vec{x} \\
+ \int \vec{u}^T \, A \, \vec{u} \, d\vec{x} + \int \tr\{(\nabla \vec{u}^T)^T B (\nabla \vec{u}^T)\} d\vec{x}
\label{eq:complete_energy}
\end{multline}
%
These regularity terms ensure that the segmenting contours in
\gls{dwi} space are still close to their native shape. The model
easily allows to incorporate inhomogeneous and anisotropic
regularization \cite{nagel_investigation_1986} to better regularize
the \gls{epi} distortion.}}

\todo[inline]{this last paragraphs in red color were located a bit
further, but I think we can keep some of the refs and still reproduce
the continuous version of the energy functional}

\section{Extensions}

\subsubsection{Semi-implicit Euler-forward optimization}
It is necessary to discretize $t$ in order to obtain a numerical
implementation of the equation \eqref{eq:gradient_final}:
\begin{align}
\frac{\vec{u}_k^{t+1}-\vec{u}_k^t}{\tau} =
&- \underset{l,m}{\sum} \underset{i}{\sum}
\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right]
\psi_k(\vec{c}_i)\, \hat{\vec{n}}_i \notag\\
&+2\, \boldsymbol{\alpha} \vec{u}_k^{t+1}
-2\, \boldsymbol{\beta} \Delta \vec{u}_k^{t+1}.
\end{align}

The associated Euler-Lagrange equation is found as:
\begin{equation}
(\tau^{-1} +2\, \boldsymbol{\alpha} - 2\, \boldsymbol{\beta} \Delta )\, \vec{u}_k^{t+1} =
\tau^{-1} \vec{u}_k^t - \frac{\partial}{\partial \vec{u}_k} E_{data}(\vec{u}_k^t),
\end{equation}
that is a linear system that we translate into Fourier domain,
to obtain the next deformation field $\vec{u}_k^{t+1}$:
\begin{multline}
\vec{u}_k^{t+1} = \\
 \mathcal{F}^{-1} \left\lbrace
\frac{\mathcal{F}\lbrace \tau^{-1}\vec{u}_{k}^{t} - \underset{l,m}{\sum} \underset{i}{\sum}
\left[ \mdist{f_i'}{l} - \mdist{f_i'}{m}\right] \rbrace}
     {\mathcal{F}\lbrace (\tau^{-1} +2\,\boldsymbol{\alpha})\mathcal{I} - 2\,\boldsymbol{\beta} \Delta ) \rbrace}
     \right\rbrace
\end{multline}

Here, we rewrite the Laplacian as a linear combination of the identity and shift operators:
\begin{equation}
\Delta = \sum\limits_d \mathcal{S}_d^- + \mathcal{S}_d^+ - 2 \mathcal{I}
\end{equation}
where $\mathcal{S}_{d}^{\pm}$ stands for the forward ($+$) and backward ($-$) shift
operator along coordinates axis $d$, of which the Fourier transform is found easily as
\begin{equation}
\mathcal{F}\{\mathcal{S}_{d}^{\pm}\} = e^{\pm \vec{i}\omega_{d}},
\end{equation}
where $\omega_{d}$ is the normalized pulsation along $d$-direction. Accordingly, the
Fourier transform of the discrete Laplacian is found as
\begin{equation}
\mathcal{F}\{\Delta\} = \sum\limits_d e^{-\vec{i}\omega_d } + e^{\vec{i}\omega_d } - 2 = \sum\limits_d \left( 2\cos(\omega_d) - 2 \right)
\end{equation}

The remaining transforms are trivial or can be computed using the \gls{fft}
as in \citep{estellers_efficient_2011}.
